Financial Contributions Made to Presidential Candidates by Aaron Gruberg
========================================================

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(tidyr)
library(dplyr)
library(readxl)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("C:/Users/Aaron/Downloads")

CA <- read.csv(file = "CA.csv", row.names = NULL)

colnames(CA) <- c(colnames(CA)[-1],"x")

# This added a dummy column named "x" that we need to remove
CA$x <- NULL 
```
# Data Description

```{r echo=FALSE, Data_Description}
# Describe the data
summary(CA)
```
> Our dataset consists of 18 variables with 1304346 observations

# Univariate Plots Section

```{r echo=FALSE, warning=FALSE, Histograms}
# Histogram of the number of people who gave each amount

ggplot(aes(x = contb_receipt_amt), data = CA) +
  geom_histogram(bins = 250) +
  scale_x_continuous(limits = c(0, 5000))


# Histogram of the log of the number of people who gave each amount

ggplot(aes(x = contb_receipt_amt), data = CA) +
  geom_histogram(bins = 250) +
  scale_x_continuous(limits = c(0, 5000)) +
  scale_y_log10()
```

> The count of contribution amounts was very spread out.  There appear to be large gaps in the first histogram between contributions of $500 and $3000.  Once I applied a log10() scale to the y axis these smaller counts were visible. 

```{r echo=FALSE, warning=FALSE, not_employed}
# Plot of contributions top candidates received from unemployed people

ggplot(data = subset(CA, CA$contbr_employer == 'NOT EMPLOYED'
                      & CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\''
                     | CA$cand_nm == 'Carson, Benjamin S.'
                     | CA$cand_nm == 'Rubio, Marco'), aes(x = cand_nm)) +
  geom_histogram(stat = "count") 
```

> This is a histogram of the number of contributions top candidates received from unemployed people.  Bernie Sanders received the most contributions from this group of people.

```{r echo=FALSE, warning=FALSE, Univariate_Plots}
# Plot of contributions top candidates received from employed people

ggplot(data = subset(CA, (CA$contbr_employer != 'NOT EMPLOYED' 
                     | CA$contbr_employer != 'RETIRED')
                     & (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\''
                     | CA$cand_nm == 'Carson, Benjamin S.'
                     | CA$cand_nm == 'Rubio, Marco')), aes(x = cand_nm)) +
  geom_histogram(stat = "count") 
```

> This is the number of contributions top candidates received from employed people.  Hillary Clinton received the most contributions from employed people.

```{r echo=FALSE, warning=FALSE, frequency}

### Amount vs Date
### Change the date format by adding new date column
CA$contb_receipt_dt_date <- as.Date(CA$contb_receipt_dt, "%d-%b-%y")

### Add new columns for Year, Day, Month
CA$year = as.numeric(format(CA$contb_receipt_dt_date, format = "%Y"))
CA$month = as.numeric(format(CA$contb_receipt_dt_date, format = "%m"))
CA$day = as.numeric(format(CA$contb_receipt_dt_date, format = "%d"))

# Frequency Plot of Total Contributions vs Date, grouped by candidate name.
ggplot(subset(CA, (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\'')),  aes(contb_receipt_dt_date, colour = cand_nm)) +
  geom_freqpoly(binwidth = 20)

```

> This is a frequency plot of the number of contributions received by date, grouped by candidate name.  Donald Trump did not receive many contributions until very close to the election date.

```{r echo=FALSE, warning=FALSE, summary}
# Summary of the number of contributions made by each profession
summary(CA$contbr_employer)
```

# Univariate Analysis

### What is the structure of your dataset?

> There are 1304346 transactions in the dataset with 18 features(cmte_id, cand_id, cand_nm, contbr_name, contbr_city, contbr_state, contbr_zip, contbr_employer, contbr_occupation, contb_receipt_amt, contb_receipt_dt, receipt_desc, memo_cd, memo_text, form_tp, file_num, tran_id, election_tp).  Most variables in the original dataset were categorical.

### What is/are the main feature(s) of interest in your dataset?

> The main features in the dataset are contb_receipt_amt and cand_nm.  I'm interested in the amount of money going to a particular candidate and how it varies across categorical variables such as employment status.  

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

> I also examine contb_receipt_dt to see how often a particular contribution is made as we get closer to the election date.

### Did you create any new variables from existing variables in the dataset?

> I created a variable called contb_receipt_dt_date which converts the date format from '26-APR-16' to '2016-04-26'.  Then I added 3 variables for the 'day', 'month', 'year'.

```{r echo=TRUE, Date_Format}
### Change the date format by adding new date column
CA$contb_receipt_dt_date <- as.Date(CA$contb_receipt_dt, "%d-%b-%y")

### Add new columns for Year, Day, Month
CA$year = as.numeric(format(CA$contb_receipt_dt_date, format = "%Y"))
CA$month = as.numeric(format(CA$contb_receipt_dt_date, format = "%m"))
CA$day = as.numeric(format(CA$contb_receipt_dt_date, format = "%d"))
```

> I created a variable 'total_contbr' to sum all the contributions a particular candidate receives on a particuluar date, and 'max_total_contbr' to find the cumulative sum a candidate received from all contributors on a particular date.



> I created the variable 'contbr_industry' to group the 'contbr_occupation' values into BLUE COLLAR, WHITE COLLAR, MILITARY, STUDENT, UNEMPLOYED.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

> There was a sharp cutoff in the histogram of contribution amounts for values of $2700.  After this limit, far fewer counts were recorded.  I was expecting to see a distribition that resembled an exponential decay but there were also large spikes for contributions of 1000, 2000, and 2700.  

> I used 'as.numeric' to convert the date values to numeric date values.  This allowed me to take the log of numeric date values for scaling purposes.

# Bivariate Plots Section


```{r echo=FALSE, Bivariate_Plots}
## Plot Matrix
# Load additional libraries
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
set.seed(20022012)
new_CA <- subset(CA, 
                 select=c(cand_nm, contb_receipt_amt, contb_receipt_dt, form_tp, file_num))

CA_sample <- new_CA[sample(1:length(new_CA$contb_receipt_amt), 10000), ]

View(head(new_CA, 200))

# Make ggpairs plot, need to increase cardinality threshold parameter!?
ggpairs(CA_sample, cardinality_threshold = 750, axisLabels = c("show", "internal", "none"),
        lower = list(continuous = wrap("points", shape = I('.')), 
                     combo = "facethist", discrete = "facetbar", na = "na"),
        upper = list(continuous = wrap("cor"), combo = "box_no_facet", 
                  discrete = "facetbar", na = "na"),
        diag = list(continuous = "densityDiag", discrete = "barDiag", na = "naDiag"))

# SCATTER PLOT OF AMOUNT VS FILENUM
ggplot(aes(x = file_num, y = contb_receipt_amt), 
       data = subset(CA, (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\''
                     | CA$cand_nm == 'Carson, Benjamin S.'
                     | CA$cand_nm == 'Rubio, Marco')
                     & CA$contb_receipt_amt > 0)) +
  geom_point()

# Make scatter plot of amount vs date
ggplot(aes(x = contb_receipt_dt_date, y = contb_receipt_amt), 
       data = CA) +
  geom_point() 

# amount vs date
# Only included dates in 2015-2016
# Only included amounts between -3000(refunds) and 3000.
ggplot(aes(x = contb_receipt_dt_date, y = contb_receipt_amt), 
       data = subset(CA, CA$year >= 2015 &
                       (CA$contb_receipt_amt > 0) &
                       (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\''
                     | CA$cand_nm == 'Carson, Benjamin S.'
                     | CA$cand_nm == 'Rubio, Marco'))) +
  geom_point() +
  scale_y_continuous(limits = c(0, 10000))

# Box Plot of amount vs cand_nm
# We are looking at how categorical variables change with candidate names.
ggplot(aes(x = cand_nm, y = contb_receipt_amt), 
       data = subset(CA, (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\''
                     | CA$cand_nm == 'Carson, Benjamin S.'
                     | CA$cand_nm == 'Rubio, Marco'))) +
  geom_boxplot() +
  scale_y_continuous(limits = c(0, 150))

# Box Plot of Amount vs Occupation
# I chose a few specific occupations
ggplot(aes(x = contbr_occupation, y = contb_receipt_amt), 
       data = subset(CA, CA$contbr_occupation == 'LAWYER'
                     | CA$contbr_occupation == 'ATTORNEY'
                     | CA$contbr_occupation == 'ENGINEER'
                     | CA$contbr_occupation == 'SOFTWARE ENGINEER'
                     | CA$contbr_occupation == 'DOCTOR'
                     | CA$contbr_occupation == 'TEACHER')) +
  geom_boxplot() +
  scale_y_continuous(limits = c(0, 400))

# Box Plot of Amount vs Employer
# for a few employers
ggplot(aes(x = contbr_employer, y = contb_receipt_amt), 
       data = subset(CA, CA$contbr_employer == 'APPLE INC.'
                     | CA$contbr_employer == 'GOOGLE'
                     | CA$contbr_employer == 'FACEBOOK'
                     | CA$contbr_employer == 'WELLS FARGO'
                     | CA$contbr_employer == 'GENENTECH'
                     | CA$contbr_employer == 'ORACLE CORP.'
                     | CA$contbr_employer == 'AT&T'
                     | CA$contbr_employer == 'NORTHROP GRUMMAN'
                     | CA$contbr_employer == 'UNITED AIRLINES'
                     | CA$contbr_employer == 'BANK OF AMERICA'
                     | CA$contbr_employer == 'VISA, INC.'
                     | CA$contbr_employer == 'DISNEY'
                     | CA$contbr_employer == 'ORACLE CORP.'
                     | CA$contbr_employer == 'WARNER BROS.'
                     | CA$contbr_employer == 'U.S. NAVY'
                     | CA$contbr_employer == '20TH CENTURY FOX')) +
  geom_boxplot() +
  scale_y_continuous(limits = c(0, 400))

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

### Add column for cumulative sum of amount, group by cand name and date
### This gives me the total contribution SO FAR, for each candidate on each date
CA$total_contbr <- ave(CA$contb_receipt_amt, CA$cand_nm, CA$contb_receipt_dt_date, FUN=cumsum)

### Add column for max of total contribution
### This gives me the MAX of total contribution for each candidate on each date
CA$max_total_contbr <- ave(CA$total_contbr, CA$cand_nm, CA$contb_receipt_dt_date, FUN=max)

# Test
#View(subset(CA, CA$year == 2015 & CA$month == 11 & CA$day == 1))

# Line Plot of Max Total Contributions vs Date, grouped by candidate name.
ggplot(subset(CA, (CA$contb_receipt_amt > 0) & 
                (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\'')),  aes(contb_receipt_dt_date, max_total_contbr, colour = cand_nm)) +
  geom_line() 

# SMOOTH LINE Plot of Max Total Contributions vs Date, grouped by candidate name.
ggplot(subset(CA, (CA$max_total_contbr > 0) & 
                (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\'')),  aes(contb_receipt_dt_date, max_total_contbr, colour = cand_nm)) +
  stat_smooth(formula = y ~ poly(x,3)) +
  scale_y_continuous(limits = c(0, max(CA$max_total_contbr)))

### Line plot of max contributions vs date, grouped by job title
### Create vectors for each 'job field' for selected jobs
WHITE_COLLAR <- c("ENGINEER", "SOFTWARE ENGINEER", "SOFTWARE DEVELOPER", "WEB DEVELOPER",
              "IT", "ARCHITECT", "SCIENTIST", "RESEARCHER", "PHYSICIAN", "REGISTERED NURSE",               "NURSE", "RN", "PHARMACIST", "NURSE PRACTITIONER", "DENTIST", "DOCTOR", "MD",                "PSYCHOLOGIST", "LAWYER", "ATTORNEY", "PARALEGAL", "JUDGE ADVOCATE", "JUDGE",
              "SALES", "SALES MANAGER", "ANALYST", "MARKETING", "CONSULTANT", "REALTOR",
              "DESIGNER", "PROJECT MANAGER", "ADMINISTRATOR", "REAL ESTATE",
              "HUMAN RESOURCES", "PROPERTY MANAGER", "REAL ESTATE BROKER", "ACCOUNTANT",
              "FINANCE", "CPA", "INVESTOR", "BANKER", "CEO", "CFO", "PRESIDENT", 
              "VICE PRESIDENT", "EXECUTIVE", "TEACHER", "PROFESSOR", "COLLEGE PROFESSOR",                  "EDUCATOR", "BEHAVIORAL THERAPIST", "PHYSICAL THERAPIST", "LECTURER",
              "MATHEMATICS INSTRUCTOR", "BUSINESS DEVELOPMENT", "BIOLOGIST", "SOCIAL WORKER",               "DATA ENTRY", "SURGEON", "DATABASE ADMINISTRATOR", "GAME DEVELOPER",
              "CHIEF FINANCIAL OFFICER", "DATA SCIENTIST", "DEV-OPS ENGINEER",
              "AEROSPACE ENGINEER")

BLUE_COLLAR <- c("BUSINESS OWNER", "SMALL BUSINESS OWNER", "OWNER", "ENTREPRENEUR", 
            "SELF EMPLOYED", "FARMER", "TRUCK DRIVER", "DRIVER", "CONTRACTOR", "BOOKKEEPER",
            "CLERICAL", "SHOP OWNER", "HOMEMAKER", "LABORER", "DELIVERY DRIVER", "SECRETARY",
            "TSA OFFICER", "SALON/SPA", "FORKLIFT DRIVER", "DATA ENTRY", "COOK",
            "SECURITY GUARD", "ELECTRICIAN", "ASSISTANT MANAGER", "RETAIL SALES", "DEALER",
            "CUSTOMER SERVICE", "FLIGHT ATTENDANT", "FIRE FIGHTER", "CARPENTER", "RANCHER")

STUDENT <- c("STUDENT", "GRADUATE STUDENT")

UNEMPLOYED <- c("DISABLED", "NOT EMPLOYED", "RETIRED", "UNEMPLOYED BECAUSE OF OBAMACARE")

MILITARY <- c("U.S. NAVY", "UNITED STATES NAVY", "NAVY", "DEPT. OF NAVY",
              "DEPARTMENT OF NAVY", "DEPARTMENT OF THE NAVY", "U.S. DEPARTMENT OF THE NAVY",
              "NAVY FEDERAL CREDIT UNION", "USNAVY", "U.S. ARMY", "ARMY", "US ARMY",
              "U.S. DEPARTMENT OF THE ARMY", "U.S. ARMY CORPS. OF ENGINEERS",
              "DEPT OF ARMY", "U S ARMY", "U.S. COAST GUARD", "US COAST GUARD",
              "UNITED STATES COAST GUARD", "U.S. DEPARTMENT OF THE COAST GUARD",
              "DOD", "D.O.D.")

### Create 'job field' column.  If contbr_occupation is in a job field vector from above,
### add the name of the vector to 'job field' column
CA$contbr_industry <- ifelse(CA$contbr_occupation %in% WHITE_COLLAR, "WHITE COLLAR",
                      ifelse(CA$contbr_occupation %in% BLUE_COLLAR, "BLUE COLLAR",
                      ifelse(CA$contbr_occupation %in% STUDENT, "STUDENT",
                      ifelse(CA$contbr_occupation %in% UNEMPLOYED, "UNEMPLOYED",
                      ifelse(CA$contbr_employer %in% MILITARY, "MILITARY","NULL")))))
# Test
#View(head(CA, 200))

### Now I can plot max contributions vs date, grouped by contbr_industry
### The y values were very close together, I'm playing with the
### y scaling and color brewer options
ggplot(subset(CA, (CA$contb_receipt_amt > 0) & 
                (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\'') &
                (CA$contbr_industry != "NULL")),
                aes(contb_receipt_dt_date, max_total_contbr, colour = contbr_industry)) +
  stat_smooth(se = FALSE, span = 0.01) +
  scale_y_continuous(limits = c(0, max(CA$max_total_contbr))) +
  scale_colour_brewer(palette = "Spectral") +
  theme(panel.background = element_rect(fill = "grey66"),
        panel.grid.major = element_line(colour = "grey50"),
        panel.grid.minor = element_line(colour = "grey60"))

### Max contributions vs date, grouped by contbr_industry, facet wrapped by cand_nm
ggplot(subset(CA, (CA$contb_receipt_amt > 0) & 
                (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\'') &
                (CA$contbr_industry != "NULL")),
                aes(contb_receipt_dt_date, max_total_contbr, colour = contbr_industry)) +
  stat_smooth(se = FALSE, span = 0.01) +
  facet_wrap(~ cand_nm, scales = "free") +
  scale_colour_brewer(palette = "Spectral") +
  theme(panel.background = element_rect(fill = "grey66"),
        panel.grid.major = element_line(colour = "grey50"),
        panel.grid.minor = element_line(colour = "grey60"))

### Max contributions vs date, grouped by contbr_occupation for chosen occupations
ggplot(subset(CA, (CA$contb_receipt_amt > 0) & 
                (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\'') &
                (CA$contbr_occupation == 'U.S. NAVY'
                     | CA$contbr_occupation == 'CUSTOMER SERVICE'
                     | CA$contbr_occupation == 'STUDENT'
                     | CA$contbr_occupation == 'SOFTWARE ENGINEER'
                     | CA$contbr_occupation == 'RETIRED')),
                aes(contb_receipt_dt_date, max_total_contbr, colour = contbr_occupation)) +
  stat_smooth(se = FALSE, span = 0.01) +
  scale_y_continuous(limits = c(0, max(CA$max_total_contbr))) +
  scale_colour_brewer(palette = "Spectral") +
  theme(panel.background = element_rect(fill = "grey66"),
        panel.grid.major = element_line(colour = "grey50"),
        panel.grid.minor = element_line(colour = "grey60"))

### Amount vs Date, colored by candidate name
ggplot(aes(x = contb_receipt_dt_date, y = contb_receipt_amt), 
       data = subset(CA, (CA$contb_receipt_amt > 0) &
                       (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\''))) +
  geom_point(aes(colour = cand_nm)) +
  scale_y_continuous(limits = c(0, 10000)) +
  scale_colour_brewer(palette = "Spectral") +
  theme(panel.background = element_rect(fill = "grey66"),
        panel.grid.major = element_line(colour = "grey50"),
        panel.grid.minor = element_line(colour = "grey60"))

### Amount vs Date, colored by contbr_industry, facet wrapped by cand_nm
ggplot(aes(x = contb_receipt_dt_date, y = contb_receipt_amt), 
       data = subset(CA, CA$year >= 2015 &
                       (CA$contb_receipt_amt > 0) &
                       (CA$contbr_industry != "NULL") &
                       (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\''
                     | CA$cand_nm == 'Carson, Benjamin S.'
                     | CA$cand_nm == 'Rubio, Marco'))) +
  geom_point(aes(colour = contbr_industry, alpha = 0.001)) +
  scale_colour_brewer(palette = "Spectral") +
  facet_wrap(~ cand_nm, scales = "free") +
  scale_y_continuous(limits = c(0, 10000)) +
  theme(panel.background = element_rect(fill = "grey66"),
        panel.grid.major = element_line(colour = "grey50"),
        panel.grid.minor = element_line(colour = "grey60"))

#View(subset(CA, CA$contb_receipt_amt > 9000))

# Amount vs Date, colored by candidate name, facet wrapped by employer
ggplot(aes(x = contb_receipt_dt_date, y = contb_receipt_amt), 
       data = subset(CA, CA$year >= 2015 &
                       (CA$contb_receipt_amt > 0) &
                       (CA$contbr_employer == 'GOOGLE'
                     | CA$contbr_employer == 'FACEBOOK'
                     | CA$contbr_employer == 'WELLS FARGO'
                     | CA$contbr_employer == 'GENENTECH'
                     | CA$contbr_employer == 'AT&T'
                     | CA$contbr_employer == 'NORTHROP GRUMMAN'
                     | CA$contbr_employer == 'UNITED AIRLINES'
                     | CA$contbr_employer == 'BANK OF AMERICA'
                     | CA$contbr_employer == 'DISNEY'
                     | CA$contbr_employer == 'WARNER BROS.'
                     | CA$contbr_employer == 'U.S. NAVY') &
                       (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\''))) +
  geom_point(aes(colour = cand_nm)) +
  scale_colour_brewer(palette = "Spectral") +
  facet_wrap(~ contbr_employer, scales = "free") +
  scale_y_continuous(limits = c(0, 5000)) +
  theme(panel.background = element_rect(fill = "grey66"),
        panel.grid.major = element_line(colour = "grey50"),
        panel.grid.minor = element_line(colour = "grey60"))

### Exploring correlation

# scatter plot of date vs max contrib, colored by cand_nm
# As max contribution values increase, they show up on fewer dates
# and those dates are later in the time period.

# Add column that converts date to numeric value

CA$numeric_date <- as.numeric(CA$contb_receipt_dt_date)
#View(head(CA, 25))

ggplot(subset(CA, (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\''
                     | CA$cand_nm == 'Carson, Benjamin S.'
                     | CA$cand_nm == 'Rubio, Marco')),  
       aes(x = max_total_contbr, y = numeric_date, colour = cand_nm)) +
  geom_point() +
  scale_y_log10() +
  scale_colour_brewer(palette = "Spectral") +
  theme(panel.background = element_rect(fill = "grey66"),
        panel.grid.major = element_line(colour = "grey50"),
        panel.grid.minor = element_line(colour = "grey60"))


## Scatter plot of date vs max contribution, colored by contbr_industry
ggplot(subset(CA, (CA$cand_nm == 'Clinton, Hillary Rodham'
                     | CA$cand_nm == 'Sanders, Bernard'
                     | CA$cand_nm == 'Trump, Donald J.'
                     | CA$cand_nm == 'Cruz, Rafael Edward \'Ted\''
                     | CA$cand_nm == 'Carson, Benjamin S.'
                     | CA$cand_nm == 'Rubio, Marco') &
                CA$contbr_industry != 'NULL'),  
       aes(x = max_total_contbr, y = numeric_date, colour = contbr_industry)) +
  geom_point() +
  scale_y_log10() +
  scale_colour_brewer(palette = "Spectral") +
  theme(panel.background = element_rect(fill = "grey66"),
        panel.grid.major = element_line(colour = "grey50"),
        panel.grid.minor = element_line(colour = "grey60"))

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!